--
title: "Week 3 data check template: PHW253 Descriptive Epidemiology Assignment"
author: "student"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
# Show R code in output for all R chunks
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

# * Purpose: Code for Week 3 data check of the Descriptive Epi Assignment
# * Data: PH W253 Descriptive Epi Assignment line list_redacted.xlsx

##### Step 1: Copy all of the code needed to complete the Week 2 data check including: 
##### library code for all packages listed in Week 1 template,
##### importing the dataset, 
##### creating new dataset and keep only rows where MLVA is an A, A2,A3 pattern
##### creating a new variable with onset date in ymd date format


# Epidemic Curve




## Step 2: Using incidence2 package, create an epi_interval dataframe


```{r}
# epicurve using incidence2 package 
# see incidence2 package documentation for all of the possible interval options 
# https://cran.r-project.org/web/packages/incidence2/incidence2.pdf




# Plot the epi curve


## Notes on default format: 
##   1) epi curve is missing: title, label on x-axis, & footnote indicating missing data     
##   2) epi curve starts at first data point
##   3) epi curve has grid lines
```
 
 
 
```{r}
# Plot the epi curve with labels and formatting using ggplot2 package for data visualization



```

